Convergent Evolution During Local Adaptation to Patchy Landscapes

### 2014-11-25 patchy-review-responses.tex

\point{2.}{
    Section "outflux of migrant alleles": I found this section much less
    clear than others and have made suggestions for a clearer presentation
    (see ms).
}

\point{3.}{
    Section "length of shared haplotype": This is very handwavy and
    (without simulation backup) I find it hard to believe that these results
    are accurate. Since it is also not central to the paper, the section
    should maybe just be dropped.
}

%%%%%%
\point{[beginning comments in the pdf]}{
``Do you assume a Gaussian kernel? Or is this irrelevant? Which properties of the kernel are required? Only existence of a finite $sigma^2$?''
}

\reply
We've tried to make it more clear here 
\llname{rr:longtailed}
exactly what we sweep under the rug later.

@@

%%%%%%
\point{}{``This is not sufficient: the patch also needs to be large enough that p_e is (almost) reached at its center. Generally, the approximation should over-estimate the true value.''}

\reply
Good point; we have adjusted this slightly. \llname{rr:p_estab}

%%%%%%
\point{[more complete information about simulation model]}{}

\reply
We've added more description of how we did the simulations
\llname{rr:sim_details},
and don't think the choice of a bounded kernel is particularly salient
(given that it's not long-tailed, of course).

@@

%%%%%%
\point{[comments near equation \eqref{eqn:eqfreq}]}{}

\reply
Both comments are excellent. 
We have added discussion of the intuition and the restrictions on $|x|$, as suggested
\llname{rr:q_form}.

%%%%%%
\point{The outflux of migrant alleles}{
    \begin{quote}
        This section is less clear than the previous one. [...]
    I think you could devote the formal derivation to the "Methods" and argue on an intuitive level in the main text:

    First, you should clearly spell out what the approximation is. You use the trunk + outgrowth picture of the branching process. Then you assume that the mutant family will only be able to cross to the other patch if the trunk hits this new patch. And when the trunk hits, you assume that all the outgrowth branches also enter it. I.e.: Eventually, either the whole family is "in" and then stays in (as in Fig 3) or no part ever enters. This approximation should be fine if the size of the new patch is sufficiently large.

    As a consequence of this, we can just follow the trunk, which moves by simple Brownian motion with an exponential life span. This defines h(x,y) as probability that the trunk arrives at the new patch.

    Now, how can we relate $h(x,y)$ to $u(x,y)$? - As follows: Once the trunk reaches the new patch (and assuming $s_m$ selection everywhere) it will live there for an exponential time with mean $1/2s_m$. Since it "lives or dies" with its entire outgrowth family (of expected size $\E[K]$) it thus contributes to the density like $u(x,y) = h(x,y) \E[K] / (2s_m)$. This motivates Eq. 11. (no need to introduce $g(A)$ in the main text)
    \end{quote}
}

\reply
   Good suggestion; as discussed above,
   we've taken what we think is a similar approach, 
   providing a more intuitive arguemnt (in \nameref{ss:heuristics})
   and merged the section referred to here with the Methods \llname{rr:moved_argument}.

%%%%%%%%
\point{[awkward sentence]}{}

\reply
Fixed? \llname{rr:awkward}

%%%%%%%%
\point{[expected total number]}{if y is a point, this number would be 0 (its a density)}

\reply
We have added the word ``informally''. \llname{rr:density}

%%%%%%%%
\point{}{ 
    \begin{quote}
    This function is non-monotonic in $s_m$ and I think this needs to be discussed: can it really be that $\migrate$ increases with $s_m$ or is this an artifact?

    In general, we need a clearer idea about the limits of this theory. There are several things to consider:
    \begin{itemize}
        \item equation (3) for the expected mutant density is only an asymptotic. Therefore, we probably need to require $|y| \gg \sigma / \sqrt(2s_m)$.

        \item the approximation $g(A) \sim 1/2s_m$ in the appendix requires that the patch size is much larger than $1/s_m$

        \item in your arguments, you use the outflow of families from a ball with (large) radius $r_0$ [although I do not see how this figures in]

        \item over short ranges, results must depend on details of the dispersal kernel anyway. [Depending on the shape of the distribution - heavy tail - this even holds for large distances]. As mentioned above: spell out assumptions on the kernel.
    \end{itemize}

    So: what is the minimal distance $|y|$ for the result to be valid? And: does this minimal distance eliminate the anomaly of an increasing $\migrate$ with $s_m$ for small $s_m?$ Alternatively: is there also a lower limit for $s_m$? I have noted that you do assume above that $s_m$ is large enough to avoid local fixation - but I'm not sure that this comes down to the same thing. 
\end{quote}
}

%%%%%%%%
\point{[expressio nfor the outflux of migrants]}{do we need this for anything?}

\reply
We don't need it for anything else, 
but we think it adds sufficiently to the intuition about the process
that we'd like to keep it in.
In practice it could provide a good reality check.
\plr{alternatively, compute this for the mice example?}

%%%%%% p.12
\point{[justify $\E[K]\approx 1/s_m$]}{}

\reply
See \eqref{eqn:mean_K}, and modified discussion around \llname{rr:EK}.

### 2014-11-25 patchy-selection-paper.tex
reviews...

%%%% %%%%
\subsubsection{Heuristics}

At first it is tempting to argue
that since the probability of establishment of a beneficial allele is $p_e$,
and in a patch of area $A$ there are $\rho A q(x)$ beneficial alleles present at equilibrium,
the mean rate of establishment is $p_e \rho A q(x)$.
This is wrong, since it ignores the family structure of migrants.

However, this wrong start suggests a correct one:
notice that the mean density $q(x)$ counts up each family's total size,
while $\migrate(x)$ counts the probability the family establishes.
We argue below that

%%%% %%%%
\subsubsection{The genealogy of migrant families}

@@

This suggests that families can be well-described by motion of a central ``trunk'',
which dies at a random time whose distribution is given by $k_e(t)$,
and conditional on survival of this trunk, the family is composed of side ``branches'' off of this trunk,
as depicted in Figure~\ref{fig:branching_decomp}.
We can understand this quite concretely, by a method described rigorously by \citet{geiger1999elementary} (see also \citet{chauvin1991growing}).
When we condition on survival until $t$, we condition on existence of
at least one trunk lineage surviving from time $0$ to time $t$ (the long red
lineage in Figure \ref{fig:branching_decomp}).
Given this lineage, nothing distinguishes any of the other reproduction events --
each individual along the trunk lineage must give birth to at least one offspring,
and all other reproduction events are unconditioned.
In other words, the genealogy of a family that survives until $t$
can be decomposed into a trunk that lasts until $t$
and that sprouts independent branches that may or may not survive until $t$.
Specifically, if we write $Z_t$ for the number of offspring in the branching process alive at time $t$,
% (so $\P\{Z_t = 0\} = k_e(t)$),
then this gives the following decomposition:
\begin{align}
  \left( Z_t \; \vert \; Z_t>0 \right) \deq 1 + \sum_{s=0}^t \sum_{k=1}^{W_s-1} Z^{(s,k)}_{t-s},
\end{align}
where each $Z^{(s,k)}$ is an independent copy of $Z$ 
that arose from the main branch in generation $s$,
and each $W_s$ is an independent copy of the offspring distribution, conditioned on $W_s \ge 1$.

@@

This suggests that we add in the spatial motion by following only the motion of the ``trunk''
(i.e.\ the red line in figure \ref{fig:branching_decomp}).
Its motion is approximately Brownian, with variance $\sigma^2$ per generation.
(Recall $\sigma$ is the dispersal distance.)
We can use these two ingredients, Brownian motion and a branching process, 
to compute the terms of equations \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}.

%%%%%%% %%%%%%%%%%%
\subsubsection{The outflux of migrant alleles}

We are approximating the dynamics of the focal allele in the region further away than $r_0$ from the patch
as the sum of independent migrant families, each of whose dynamics are
given by a spatial branching process (as depicted in Figure \ref{fig:branching_decomp}).
Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\partial B(r_0)$ its boundary.
Suppose that there is a constant outflux of migrant families from each point $x \in \partial B(r_0)$,
such that the long-term proportion of individuals near a point $x$ in $\partial B(r_0)$ that initiate new migrant families
is $\gamma(x)$.
To make expressions \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate} more concrete,
we need to replace ``outflux of families'' by a sum over points at distance $r_0$ from the patch
of the outflux $\gamma$ at that point.
By examining the constituent terms,
it will turn out that equation \eqref{eqn:gestalt_migrate} 
is well-approximated by a constant multiple of \eqref{eqn:gestalt_q}.

First consider equation \eqref{eqn:gestalt_q} for the equilibrium frequency.
Suppose that $Z$ is one such spatial branching process as above, started at time 0 with a single individual at $x$,
and denote by $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the spatial region $A$.
The \emph{mean occupation density} of $Z$ at location $y$
can be thought of 
as the expected total number of offspring of a family beginning at $x$ that ever live at $y$.
It is defined to be the density of $\E[\int_0^t Z_t(\cdot) dt]$,
i.e.\ the smooth function $u(x,y)$ that, integrated over a set of locations $S$, 
gives the expected total number of individuals ever living in $S$:
\begin{align} \label{eqn:occupation_density_defn}
    \E\left[ \int_0^\infty Z_t(S) dt \right] = \int_S u(x,y) dy .
\end{align}
% Said another way, $u(x,y)$ is the expected number of offspring of a progenitor at $x$ that will ever be observed near $y$.
Since reproduction and spatial motion are independent,
$u(x,y)$ is the sum across all generations of the expected number of individuals alive in that generation,
multiplied by the probability density that spatial motion takes a lineage from $x$ to $y$,
i.e.\
\begin{align}
    u(x,y) = \int_0^t e^{-s_m t} p_t(x,y) dt .
\end{align}
The equilibrium frequency as decomposed in expression \eqref{eqn:gestalt_q} 
is the sum of mean occupation densities of a constant outflux of branching processes
from $\partial B(r_0)$:
\begin{align} \label{eqn:occupation_integral}
    q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  . % \\
         % &= \int_{\partial B(r_0)} \gamma(x) \int_0^t e^{-s_m t} p_t(x,y) dt dx .
\end{align}

This form we can now compare to 
the expression \eqref{eqn:gestalt_migrate} for the outflux of successful migrants.
Consider the probability that a migrant family beginning with a single individual at $x$ will ever establish in the new patch.
We assume this new patch is circular, 
has relatively small area $A$ and is centered at position $y$.
% The probability that this family reaches $A$ at time $t$ is the probability that it survives until $t$, $(1-k_e(t))$,
% multiplied by the probability that the family reaches the patch, given survival.
% If $t$ is large then as in the previous section, $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$.
It would be possible to analyze this probability directly,
as in \citet{barton1987establishment};
but here we take a simpler route, 
approximating this by the chance that the trunk hits the new patch,
multiplied by the chance that at least one member of the family escapes demographic stochasticity
and successfully establishes in the new patch.
Write $h_A(x,y)$ for the probability that the Brownian trunk
hits the new patch centered at $y$ before the family dies out,
and $f(A)$ for the chance that the family manages to establish in the new patch,
given that it successfully arrives.
As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is properly an integral against the outflux $\gamma(x)$:
\begin{align}\label{eqn:migrate_integral}
  \migrate(y) &= \rho \int_{\partial B(r_0)} \gamma(x) h_A(x,y) f(A) dx .
\end{align}

Intuitively, $h_A(x,y)$ counts each family once if it hits the patch,
while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
Since the expected total time spent in the patch does not depend on the distance at which the family started
if we know that it survived to hit the patch,
it is intuitively reasonable that $u(x,y)$ and $h_A(x,y)$ are simply related.
Following this intuition,
and that $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
it turns out (see section~\nameref{apx:hitting_occupation}) that
\begin{align} 
    h_A(x,y) \approx \frac{ u(x,y) }{ \E[K] g(A) } ,
\end{align}
where $g(A)$ is the total expected occupation time of the patch given the family arrives.
Recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
Since the integrands in expressions \ref{eqn:occupation_integral} and \ref{eqn:migrate_integral} only differ by a factor
that (at least asymptotically) does not depend on the distance between $x$ and $y$,
we can obtain the migration rate by multiplying the equilibrium frequency by this factor.
The probability $f(A)$ that the family does not die out once it arrives at the new patch is
approximately $\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$,
if each member arriving has chance $p_e$ of establishing, independently.
\plr{And if $p_e$ is small?}
% If we let $w$ be the width of the patch, then the remaining term is
% \begin{align}
%   s(w) \simeq \frac{1}{2s_m} \left( 1 - \left( \frac{\pi w \sqrt{2 s_m}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2s_m}/\sigma} \right).
% \end{align}
Combining with the fact that $g(A)\approx 1/(2s_m)$ (see section~\nameref{apx:hitting_occupation}, in particular equation \eqref{eqn:q_migrate_relation}),
we obtain that
\begin{align}
  \label{eqn:migrate} 
  \migrate(y) &\approx q(y) \times \frac{ 4 \rho s_p  s_m }{ \xi^2 }  \\
  &\approx C \frac{ 4 \rho s_p s_m }{ \xi^2 } \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma) .
\end{align}
This does not depend on $A$, because although
$f(A)$ and $g(A)$ in principle depend on the patch size (and geometry),
the dependence is very weak.
For instance, the width of a circular patch only very weakly affects the chance 
of establishment of a new migrant that appears on its edge,
as long as the patch is large enough.
For more discussion, including an expression for the outflux of migrants $\gamma(x)$, see section~\nameref{apx:hitting_occupation}.

@@

Our first obstacle is that we do not know $1-k_e(t)$, 
the probability a family survives until $t$, exactly.
However, we do know that $1-k_e(t) \simeq e^{-s_m t} / \E[K]$,
i.e.\ that it has exponential tails.
Since $\E[K] \approx 1/s_m$, it is sensible to make one further approximation,
supposing that the random lifetime of a family is exponentially distributed with mean $1/s_m$.
Once we do this, 
we can use standard results on hitting times of $d$-dimensional Brownian motion
that is killed at rate $s_m$ (see \citet{borodin2002handbook} 2.2.0.1 and 4.2.0.1).
In particular, if the patch is circular with radius $w$ and lies at distance
$R$ from the already adapted patch, then 
we know that the Laplace transform of the transit time $\tau$ is
\begin{align}
  \E[e^{-\ell \tau}] &=
    \begin{cases}
      e^{- R \sqrt{2(s_m+\ell)}/\sigma} \qquad & d=1 \\
      \frac{ K_0( (R+w)\sqrt{2(s_m+\ell)}/\sigma) }{ K_0( w\sqrt{2(s_m+\ell)}/\sigma) } \qquad & d=2  ,
    \end{cases} \label{eqn:borodinresult}
    % &= \left( R/w +1 \right)^{-\frac{d-1}{2}} e^{- R \sqrt{2(s_m+\ell)}/\sigma}  (1+O(1/w)) .
\end{align}
where $K_0$ is a modified Bessel function of the second kind.
We are interested in lineages that manage to reach the patch before being killed,
i.e.\ having $\tau < \infty$,
which occurs with probability
$\P\{\tau < \infty\} = \lim_{\ell \to 0} \E \left[\exp(-\ell \tau) \right]$.

@@

%%%%%%%%% %%%%%%%%%%%%%%
\subsection[Hitting and Occupation]{Hitting and occupation}
\label{apx:hitting_occupation}

>>>>

%%%% %%%%
\subsubsection{The genealogy of migrant families}
\label{ss:migrant_math}

@@

This suggests that families can be well-described by motion of a central ``trunk'',
which dies at a random time whose distribution is given by $k_e(t)$,
and conditional on survival of this trunk, the family is composed of side ``branches'' off of this trunk,
as depicted in Figure~\ref{fig:branching_decomp}.
We can understand this quite concretely, by a method described rigorously by \citet{geiger1999elementary} (see also \citet{chauvin1991growing}).
When we condition on survival until $t$, we condition on existence of
at least one trunk lineage surviving from time $0$ to time $t$ (the long red
lineage in Figure \ref{fig:branching_decomp}).
Given this lineage, nothing distinguishes any of the other reproduction events --
each individual along the trunk lineage must give birth to at least one offspring,
and all other reproduction events are unconditioned.
In other words, the genealogy of a family that survives until $t$
can be decomposed into a trunk that lasts until $t$
and that sprouts independent branches that may or may not survive until $t$.
Specifically, if we write $Z_t$ for the number of offspring in the branching process alive at time $t$,
% (so $\P\{Z_t = 0\} = k_e(t)$),
then this gives the following decomposition:
\begin{align} \label{eqn:Zt_decomposition}
  \left( Z_t \; \vert \; Z_t>0 \right) \deq 1 + \sum_{s=0}^t \sum_{k=1}^{W_s-1} Z^{(s,k)}_{t-s},
\end{align}
where each $Z^{(s,k)}$ is an independent copy of $Z$ 
that arose from the main branch in generation $s$,
and each $W_s$ is an independent copy of the offspring distribution $W$, conditioned on $W_s \ge 1$.
Since each subfamily $Z^{(s,k)}$ is a branching process with mean offspring number less than one,
most of these die out quickly,
and so $Z_t$ is composed of a cluster of recently split subfamilies.
If we take the expectation of \eqref{eqn:Zt_decomposition},
we get the limiting mean family size conditioned on nonextinction:
\begin{align} \label{eqn:mean_K}
    \E[K] = 1 + \E[W \given W \ge 1] /(e^{s_m}-1) \approx \E[W \given W \ge 1] / s_m .
\end{align}

@@

%%%% %%%%
\subsubsection{Heuristics}
\label{ss:heuristics}

The parallel form suggested by \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}
suggests that we can find $\migrate(x)$ by multiplying the mean density $q(x)$
by the ratio of the second terms in the two equations,
i.e.
\begin{align} \label{eqn:ratio_gestalt}
  \migrate(x) 
    &= \rho q(x) \times 
    \frac{
        \text{ ( probability a family establishes in patch at $x$ ) }
    }{
        \text{ ( mean occupation density of a family at $x$ ) }
    } .
\end{align}
To make this precise we need to understand what we mean by ``outflux of families'';
we do this below in \nameref{ss:hitting_occupation},
and provide a heuristic argument here.

Both numerator and denominator in \eqref{eqn:ratio_gestalt} 
count only families that have arrived at the new patch,
so we can consider a family conditioned on having successfully transited between the patches.
To obtain simply interpretable expressions,
we assume that once any member of the family arrives at the patch,
all other members are close by, and so have roughly equal opportunity
to establish in the new patch.

Suppose the family arrives with $K$ individuals,
i.e.\ there are $K$ individuals that manage to establish
The probability that at least one individual leaves a lineage that establishes in the patch
is $1-(1-p_e)^K$.
On the other hand, in the absence of the patch of new habitat,
each of the $K$ newly arrived individuals would leave behind a lineage with 
mean total size $\int_0^\infty e^{-t s_m} dt = 1/s_m$,
roughly half of which would stay in the new patch.
Therefore, the expected contribution of the family to the occupation density is $K/2s_m$.
(Readers who feel uncomfortable about the geometry of the patch,
or the distinction between occupation density and number of individuals
are encouraged to take a detour through \nameref{ss:hitting_occupation}, below.)

This suggests that 
\begin{align} \label{eqn:ratio_K}
  \migrate(x) 
    &= \rho q(x) \times 
    \frac{
        \E[1-(1-p_e)^K]
    }{
        \E[K]/2s_m
    } .
\end{align}
There are now two extremes to consider.
If $\P\{K>1/p_e}$ is small, then each family has a small chance of establishment,
and the numerator is approximately $p_e \E[K]$.
On the other hand, if $p_e$ is not too small,
then the numerator would be close to 1, as any family that arrives will likely establish.
Combining these, and using that $\E[K] \approx 1/s_m$, 
\begin{align} \label{eqn:twocase_migrate}
    \migrate(x) &\approx 2 \rho s_m \min(s_m,p_e) \times q(x) .
%     \begin{cases}
%         q(x) \times 4 \rho s_m s_p / \xi^2  \qquad & \text{ if } s_p < s_m \\
%         q(x) \times 2 \rho s_m^2 \qquad & \text{ if } s_p > s_m 
%     \end{cases} .
\end{align}
When we use this, we set $p_e = 2 s_p / \xi^2$.

@@

Our first obstacle is that we do not know $1-k_e(t)$, 
the probability a family survives until $t$, exactly.
However, we do know that $1-k_e(t) \simeq e^{-s_m t} / \E[K]$,
i.e.\ that it has exponential tails.
This suggests one further approximation, 
that the random lifetime of a family is exponentially distributed with mean $1/s_m$;
equation \eqref{eqn:mean_K} shows this introduces a constant into the equations below. \linelabel{rr:EK}
Once we do this, 
we can use standard results on hitting times of $d$-dimensional Brownian motion
that is killed at rate $s_m$ (see \citet{borodin2002handbook} 2.2.0.1 and 4.2.0.1).
In particular, if the patch is circular with radius $w$ and lies at distance
$R$ from the already adapted patch, then 
we know that the Laplace transform of the transit time $\tau$ is
\begin{align}
  \E[e^{-\ell \tau}] &=
    \begin{cases}
      e^{- R \sqrt{2(s_m+\ell)}/\sigma} \qquad & d=1 \\
      \frac{ K_0( (R+w)\sqrt{2(s_m+\ell)}/\sigma) }{ K_0( w\sqrt{2(s_m+\ell)}/\sigma) } \qquad & d=2  ,
    \end{cases} \label{eqn:borodinresult}
    % &= \left( R/w +1 \right)^{-\frac{d-1}{2}} e^{- R \sqrt{2(s_m+\ell)}/\sigma}  (1+O(1/w)) .
\end{align}
where $K_0$ is a modified Bessel function of the second kind.
We are interested in lineages that manage to reach the patch before being killed,
i.e.\ having $\tau < \infty$,
which occurs with probability
$\P\{\tau < \infty\} = \lim_{\ell \to 0} \E \left[\exp(-\ell \tau) \right]$.

@@

%%%%%%%%% %%%%%%%%%%%%%%
\subsection[Hitting and Occupation]{Hitting and occupation}
\label{ss:hitting_occupation}

Here we make a more precise argument to back up expression \eqref{eqn:twocase_migrate} for the migration rate.
The argument made above in \nameref{ss:heuristics} applies to general migration mechanisms,
since it relies only on a decomposition of the migrant families upon hitting the new patch;
but it is also imprecise in a number of subtle ways
that are difficult to formalize.
The argument we make here therefore takes a somewhat different tack,
supposing that it suffices to model the spatial movement of a migrant family
by following only the motion of the ``trunk''
(i.e.\ the red line in figure \ref{fig:branching_decomp}),
and supposing this motion is Brownian, with variance $\sigma^2$ per generation.
(Recall $\sigma$ is the dispersal distance.)
We then use facts about Brownian motion and branching processes, 
to compute more precise versions of equations \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate}.
\linelabel{rr:moved_argument}

We are approximating the dynamics of the focal allele in the region further away than $r_0$ from the patch
as the sum of independent migrant families, each of whose dynamics are
given by a spatial branching process (as depicted in Figure \ref{fig:branching_decomp}).
Call $B(r_0)$ the region closer than $r_0$ to the patch, and $\partial B(r_0)$ its boundary.
Suppose that there is a constant outflux of migrant families from each point $x \in \partial B(r_0)$,
such that the long-term proportion of individuals near a point $x$ in $\partial B(r_0)$ 
that initiate new migrant families is $\gamma(x)$.
Now note that expressions \eqref{eqn:gestalt_q} and \eqref{eqn:gestalt_migrate} are a simple product of the ``ouflux of families'',
when in fact they should be an integral of $\gamma(x)$ over possible locations. \linelabel{rr:awkward}
By examining the constituent terms,
it will turn out that equation \eqref{eqn:gestalt_migrate} 
is well-approximated by a constant multiple of \eqref{eqn:gestalt_q}.

First consider equation \eqref{eqn:gestalt_q} for the equilibrium frequency.
Suppose that $Z$ is one such spatial branching process as above, started at time 0 with a single individual at $x$,
and denote by $Z_t(A)$ is the number of descendant individuals alive at time $t$ living in the spatial region $A$.
The \emph{mean occupation density} of $Z$ at location $y$
can be thought of  informally \linelabel{rr:density}
as the expected total number of offspring of a family beginning at $x$ that ever live at $y$.
It is defined to be the density of $\E[\int_0^t Z_t(\cdot) dt]$,
i.e.\ the smooth function $u(x,y)$ that, integrated over a set of locations $S$, 
gives the expected total number of individuals ever living in $S$:
\begin{align} \label{eqn:occupation_density_defn}
    \E\left[ \int_0^\infty Z_t(S) dt \right] = \int_S u(x,y) dy .
\end{align}
% Said another way, $u(x,y)$ is the expected number of offspring of a progenitor at $x$ that will ever be observed near $y$.
Since reproduction and spatial motion are independent,
$u(x,y)$ is the sum across all generations of the expected number of individuals alive in that generation,
multiplied by the probability density that spatial motion takes a lineage from $x$ to $y$,
i.e.\
\begin{align}
    u(x,y) = \int_0^t e^{-s_m t} p_t(x,y) dt .
\end{align}
The equilibrium frequency as decomposed in expression \eqref{eqn:gestalt_q} 
is the sum of mean occupation densities of a constant outflux of branching processes
from $\partial B(r_0)$:
\begin{align} \label{eqn:occupation_integral}
    q(y) &= \int_{\partial B(r_0)} \gamma(x) u(x,y) dx  . % \\
         % &= \int_{\partial B(r_0)} \gamma(x) \int_0^t e^{-s_m t} p_t(x,y) dt dx .
\end{align}

This form we can now compare to 
the expression \eqref{eqn:gestalt_migrate} for the outflux of successful migrants.
Consider the probability that a migrant family beginning with a single individual at $x$ will ever establish in the new patch.
We assume this new patch is circular, 
has relatively small area $A$ and is centered at position $y$.
% The probability that this family reaches $A$ at time $t$ is the probability that it survives until $t$, $(1-k_e(t))$,
% multiplied by the probability that the family reaches the patch, given survival.
% If $t$ is large then as in the previous section, $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$.
It would be possible to analyze this probability directly,
as in \citet{barton1987establishment};
but here we take a simpler route, 
approximating this by the chance that the trunk hits the new patch,
multiplied by the chance that at least one member of the family escapes demographic stochasticity
and successfully establishes in the new patch.
Write $h_A(x,y)$ for the probability that the Brownian trunk
hits the new patch centered at $y$ before the family dies out,
and $f(A)$ for the chance that the family manages to establish in the new patch,
given that it successfully arrives.
As for $q(y)$ above, expression \eqref{eqn:gestalt_migrate} is properly an integral against the outflux $\gamma(x)$:
\begin{align}\label{eqn:migrate_integral}
  \migrate(y) &= \rho \int_{\partial B(r_0)} \gamma(x) h_A(x,y) f(A) dx .
\end{align}
Recall that $\rho$ appears here but not in \eqref{eqn:occupation_integral}
because $q(x)$ is the local \emph{proportion} of $B$ alleles.

Intuitively, $h_A(x,y)$ counts each family once if it hits the patch,
while the occupation density $u(x,y)$ counts the total amount of time that the family spends in the patch.
Since the expected total time spent in the patch does not depend on the distance at which the family started
if we know that it survived to hit the patch,
it is intuitively reasonable that $u(x,y)$ and $h_A(x,y)$ are simply related.
Following this intuition,
and that $(1-k_e(t)) \approx e^{-s_m t}/\E[K]$,
it turns out (see section~\nameref{ss:hitting_occupation}) that
\begin{align} 
    h_A(x,y) \approx \frac{ u(x,y) }{ \E[K] g(A) } ,
\end{align}
where $g(A)$ is the total expected occupation time of the patch given the family arrives.
Recall that $\E[K]$ is the expected family size after a large number of generations, given survival.
Since the integrands in expressions \ref{eqn:occupation_integral} and \ref{eqn:migrate_integral} only differ by a factor
that (at least asymptotically) does not depend on the distance between $x$ and $y$,
we can obtain the migration rate by multiplying the equilibrium frequency by this factor.
The probability $f(A)$ that the family does not die out once it arrives at the new patch is
approximately $\E[1-(1-p_e)^K] \approx 2 s_p \E[K] / \xi^2$,
if each member arriving has chance $p_e$ of establishing, independently.
\plr{And if $p_e$ is small?}
% If we let $w$ be the width of the patch, then the remaining term is
% \begin{align}
%   s(w) \simeq \frac{1}{2s_m} \left( 1 - \left( \frac{\pi w \sqrt{2 s_m}}{ 2 \sigma } \right)^{\frac{d-1}{2}} e^{-w \sqrt{2s_m}/\sigma} \right).
% \end{align}
Combining with the fact that $g(A)\approx 1/(2s_m)$ (see section~\nameref{ss:hitting_occupation}, in particular equation \eqref{eqn:q_migrate_relation}),
we obtain that
\begin{align}
  \label{eqn:migrate} 
  \migrate(y) &\approx q(y) \times \frac{ 4 \rho s_p  s_m }{ \xi^2 }  \\
  &\approx C \frac{ 4 \rho s_p s_m }{ \xi^2 } \left( |y| \sqrt{2 s_m}/\sigma \right)^{-(d-1)/2} \exp( - |y| \sqrt{2 s_m} / \sigma) .
\end{align}
This does not depend on $A$, because although
$f(A)$ and $g(A)$ in principle depend on the patch size (and geometry),
the dependence is very weak.
For instance, the width of a circular patch only very weakly affects the chance 
of establishment of a new migrant that appears on its edge,
as long as the patch is large enough.
For more discussion, including an expression for the outflux of migrants $\gamma(x)$, see section~\nameref{ss:hitting_occupation}.
